\par\bigskip  
\par\bigskip

During the last few years I have had the opportunity to work and share fun times with a lot 
of brilliant and kind people, and this has really been at the heart of my PhD. 

\par\medskip 

% Aurelio
First of all, I would like to thank my PhD supervisor, Aurelio Juste, who has always been 
extremely supportive of my work; 
he has let me free to explore my ideas, while being constantly available to provide guidance and advice. 
Aurelio and Martine Bosman have been able to create a great working environment, and 
all the people who are (or have been) part of the IFAE group have been precious colleagues and 
good friends.  
All the post-docs in the group have always been available to help me with 
anything from technicalities to long physics discussions, 
and I am particularly indebted to Lo\"ic Valery and Trisha Farooque, who have helped me growing as a scientist and have offered me 
their friendship. 
Carlos, Davide, Julian, Mirko, Nicola and Tal have been great office mates 
and a pleasure to spend the time with. 
 % for the past four years.
 
\par\medskip 

% Tommaso
This PhD has not been the beginning of my experience in ATLAS, and I am extremely grateful to 
Tommaso Lari, who supervised me during my bachelor and master thesis at the University of Milan, 
and continued being a reference figure also during my PhD; 
he taught me how to appreciate the most interesting aspects of High Energy Physics research 
and how to have fun while doing it.   

\par\medskip 

% multi-b team
Since the beginning of the PhD I have found in the multi-$b$ team a 
group of people who kept me motivated and helped me 
gaining independence in my work, 
in an atmosphere that always enhanced collaboration. % and friendship.  
I am particularly grateful to Max Swiatlowski, whose expertise, energy and dedication have been 
invaluable ingredients in building a wonderful team, 
and to Giordon Stark, who has always been available to patiently help me and many others (and 
to whom goes also the credit for the awesome event display on the cover). 

\par\medskip
% SUSY conveners and sub-conveners 
I have had the chance to work with many great people in the ATLAS SUSY group, and 
I want to thank particularly Iacopo Vivarelli, Till Eifert and Zach Marshall, who have been conveners of the group,
and all the subconveners who in the past years 
have put a lot of effort and passion in organizing the group and providing support and critical review 
to each analysis. 

\par\medskip 
I would like to thank the people working in TileCal luminosity group, 
and I am especially grateful to 
Ilya Korolkov, Richard Hawkings and Witold Kozanecki for their patience and their inspiring ideas. 
I also would like to thank all the people of the HistFitter team, 
particularly Sophio Pataraia and Jeanette Lorenz. 

\par\medskip 
% Andrea and Carlo 
Moving to a new place is always somehow scary, but since the beginning of my 
stay at CERN in 2014 I have had the opportunity to rely on Andrea, a fellow PhD student in the 
IFAE group. 
Her amazing human skills and optimism have helped me making a lot of good friends, including Dott. Ing. 
Carlo Zanoni, who have been an important part of my everyday life. 

\par\medskip

% Milano people
%People who pass by the Milano ATLAS group never really leave it. 
During the last few years a lot of people that are currently studying in Milan, or that like me 
studied in Milan to then continue working somewhere else, created a lively CERN-based community that 
makes me feel closer to home. 
Thanks to Alessandra, Carlo, Davide, Ettore, Giacomo, Giulia, Maria Giulia, Stefano and the rest of the 
lunch crew! A special thank you goes to the SUSY-enthusiast Claudia, Sonia and Lorenzo who have helped me in the last-minute panic moments. 

\par\medskip 
% Mamma papa' e Laura 
My family has been supportive of my choices for my entire life. 
Without the help and love of my mom, my dad and Laura I would not have even been able to start this work. 
Thank you for giving me all the opportunities I could ask for. 

\par\medskip 
% Topolino 
Throughout these years I have had one person who never doubted that I would make it, 
who lived with me through the darkest times of missed deadlines 
and who still makes me smile every single day independently 
of everything else. 
Thank you Giuseppe for knowing me better than anybody else, 
and deciding to share your life with me! 

%I have said thank you to a lot of people who contributed to make these last few years a memorable experience. 
%But it has not been always easy. There have been days when I was so stressed about missed deadlines, 
%about failures at multi-tasking, that 
%I would barely speak to anybody. 
%Days when I doubted I could make it. 

%I am lucky enough that I can't really say ``I would never have made it without you'', 
%because I don't know how it would have been without you, you have always been there for me. 
